% The Quantum Balance — Print (6×9" KDP)
\documentclass[12pt, twoside]{book}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm}
\usepackage{geometry}
\geometry{paperwidth=15.24cm, paperheight=22.86cm,
  inner=2.8cm, outer=1.8cm, top=2.2cm, bottom=2.5cm}
\usepackage{setspace}\onehalfspacing
\setlength{\headheight}{14.5pt}\addtolength{\topmargin}{-2.5pt}
\usepackage[protrusion=true, expansion=false]{microtype}
\usepackage{emptypage}
\usepackage{graphicx}\graphicspath{{figures/}}
\usepackage{booktabs, array, longtable, multirow}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem, xcolor}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{algorithm, algpseudocode}
\floatname{algorithm}{Algorithm}

% ─── Chapter style ───────────────────────────────────────────────────────────
\titleformat{\chapter}[display]
  {\normalfont\Large\bfseries}{\chaptertitlename~\thechapter}{12pt}{\Huge}
\titlespacing*{\chapter}{0pt}{30pt}{20pt}

\titleformat{\part}[display]
  {\centering\normalfont\huge\bfseries}{Part~\thepart}{20pt}{\Huge}

% ─── Header/footer ───────────────────────────────────────────────────────────
\pagestyle{fancy}
\fancyhf{}
\fancyhead[LE]{\small\textit{The Quantum Balance}}
\fancyhead[RO]{\small\textit{\leftmark}}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% ─── Theorems ────────────────────────────────────────────────────────────────
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{conjecture}[definition]{Conjecture}
\newtheorem{hypothesis}[definition]{Hypothesis}
\theoremstyle{remark}
\newtheorem{remark}{Remark}[chapter]
\newtheorem{openprob}{Open problem}[chapter]

% ─── Epistemic macros ────────────────────────────────────────────────────────
\newcommand{\Proved}{\textnormal{\textbf{[PROVED]}}}
\newcommand{\Hypoth}{\textnormal{\textbf{[WORKING HYPOTHESIS]}}}
\newcommand{\Conj}{\textnormal{\textbf{[CONJECTURE]}}}
\newcommand{\ToVerify}{\textnormal{\textbf{[TO VERIFY]}}}

% ─── Pandoc compatibility ────────────────────────────────────────────────────
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\newcounter{none}

% ─── Math macros ─────────────────────────────────────────────────────────────
\newcommand{\Ts}{T_s^{(2),\text{sys}}}
\newcommand{\deq}{\mathrm{DEQ}^2}

% ─── PDF metadata ────────────────────────────────────────────────────────────
\hypersetup{
  pdftitle={The Quantum Balance},
  pdfauthor={Luis Fernando Marcelino},
  pdflang={en},
  pdfsubject={Geopolitics, Complex Systems, Hegemonic Transition},
  pdfkeywords={DEQ2, hegemony, Kuramoto, phase transition, Brazil, USA, China, formal geopolitics}
}

% ─── Document ────────────────────────────────────────────────────────────────
\begin{document}

%% ── Title page ────────────────────────────────────────────────────────────
\begin{titlepage}
  \centering\vspace*{1.5cm}
  {\Huge\bfseries The Quantum\par}
  \vspace{0.2cm}
  {\Huge\bfseries Balance\par}
  \vspace{0.8cm}
  \rule{0.8\linewidth}{0.5pt}
  \vspace{0.6cm}

  {\large Geometry of Hegemony\\[0.3cm]
   and Phase Transition 2028--2031\par}
  \vspace{2cm}
  {\large Luis Fernando Marcelino\par}
  \vspace{0.5cm}
  {\normalsize Milan, 2026\par}
  \vfill
  {\small\textit{New Horizons Series} --- Volume~4\par}
  \vspace{0.5cm}
  \begin{small}\begin{quote}
    \textit{Epistemic labels used in the text:}\\[2pt]
    \Proved{} formally verified or confirmed on all examined cases;\\
    \Hypoth{} declared working hypothesis;\\
    \Conj{} observational conjecture;\\
    \ToVerify{} claim to be verified empirically.
  \end{quote}\end{small}
\end{titlepage}

%% ── Copyright page ────────────────────────────────────────────────────────
\thispagestyle{empty}
\vspace*{\fill}
\begin{small}
\noindent \textcopyright{} 2026 Luis Fernando Marcelino\\[4pt]
All rights reserved.\\[4pt]
First digital and print edition, 2026.\\[4pt]
\textit{The Quantum Balance} is the fourth volume of the \textit{New Horizons} series.\\[4pt]
Preceded by:\\
1. \textit{The Marcelino Constant} (Marcelino \& Teixeira, 2026)\\
2. \textit{The Geometry of Hegemony} (Marcelino, 2026)\\
3. \textit{MQNU --- Quantum Model of Human Nature} (Marcelino, 2026)\\[4pt]
The micro-foundation of the DEQ² model is developed in:\\
\textit{The Topography of Emancipatory Discourse} (Marcelino, 2026, 170pp).
\end{small}
\vspace*{\fill}
\clearpage

%% ── Table of Contents ──────────────────────────────────────────────────────
\pagenumbering{roman}
\tableofcontents
\cleardoublepage

%% ── Technical Preface ─────────────────────────────────────────────────────
\chapter*{Technical Preface}
\addcontentsline{toc}{chapter}{Technical Preface}
\markboth{Technical Preface}{Technical Preface}

\textbf{Position in the intellectual project.}
This volume is the fourth in the \textit{New Horizons} series and presupposes --- without requiring --- prior reading of the three previous volumes. New readers may proceed linearly: Chapters~1 and~2 synthesize the necessary apparatus from the \textit{Topography} and the \textit{Geometry}; Chapter~3 builds the bridge between the two scales. Readers familiar with the prior volumes will find a compact synthesis in the first two chapters and the original extension in the subsequent ones.

\textbf{A note on quantum mechanics as structural isomorphism.}
The title contains the word ``quantum.'' This requires prior clarification. The book does not claim that geopolitics \textit{is} quantum mechanics, nor that heads of state behave like particles. The claim is more precise and more limited: that the mathematical structure of the Kuramoto model --- phase transitions, order parameter, critical threshold, decoherence --- constitutes a \textit{structural isomorphism} with certain patterns of hegemonic dynamics. Just as Boltzmann's statistical thermodynamics does not claim that gases are ``truly'' probabilistic but that the probabilistic tool correctly describes macroscopic properties, the DEQ² uses the Kuramoto formalism as a descriptive-predictive instrument, not an ontological claim.

The epistemological reference is developed in \textit{MQNU --- Quantum Model of Human Nature} (Marcelino, 2026), particularly in Chapter~1 (structural isomorphism vs. ontological reduction) and Chapter~4 (limits of the analogy).

\textbf{A note on the evolutionary foundation of cooperation.}
The DEQ$^2$ describes \textit{where} hegemonic systems are headed --- but required, until this edition, an answer to the \textit{why}. The why is provided by evolutionary biology. Hamilton's condition (1964) --- $rB > C$ --- establishes when cooperation is evolutionarily stable in a population. The Price equation (1970) algebraically decomposes the variation in cooperative behaviour into an inter-group term (proportional to coherence $\Phi$) and an intra-group term (proportional to involution $\Omega$). Multilevel selection theory (Wilson \& Sober 1994; Okasha 2006) formally justifies the level-shift in description that the DEQ$^2$ performs. Nowak (2006) provides the diagnostic taxonomy of cooperative mechanisms; Ostrom (1990) the empirical foundation for the normative dimension.

The \textit{Captured Public Goods Game} (CPGG, Appendix~D §D.4) formalizes the mechanism by which an extractive sub-network of fraction $\phi$ simultaneously degrades all cooperative mechanisms, lowering the effective degree $k_\text{eff}(\phi)$ of the cooperative network and driving $\Phi \to 0$. From this analysis one \emph{derives} --- rather than postulates --- the correct formula $\Omega \approx \phi\eta r/(r-1)$ (Ch.~3). Monitoring $\phi$ thus becomes the primary early warning indicator of $T_s^{(2),\text{sys}} \to 0$.

Kropotkin (1902) had phenomenologically anticipated the central thesis: cooperation is the natural state, decoherence is the engineered state. The CPGG formalism now provides the rigorous derivation of that intuition.

\textbf{Structure and how to read.}
The book is divided into five parts. Parts~I--II construct the theoretical apparatus. Part~III analyzes the four structural actors of the global system 2024--2031. Part~IV presents the SPC-DEQ Dashboard and the falsifiable predictions. Part~V provides action protocols for individuals, organizations, and states.

Appendices~A--D contain complete formal derivations, empirical calibration, measurement protocols, and dialogue with relevant intellectual traditions. They are designed to be read independently by readers with a technical background.

\textbf{Falsifiability.}
This book makes dated, numerical predictions. Chapter~11 lists 16 falsifiable predictions with explicit time windows (2026--2031). The public pre-registration protocol is accessible at the GitHub repository \texttt{bilanciere-quantistico/predizioni-deq2}. If the predictions were to fail systematically, the DEQ² model would need revision.

\cleardoublepage
\pagenumbering{arabic}

%% ═══════════════════════════════════════════════════════════════════════════
%% PART I — From the Communicative Micro to the Macro Geopolitics
%% ═══════════════════════════════════════════════════════════════════════════
\part[From the Communicative Micro to Macro Geopolitics]{From the Communicative Micro\\to Macro Geopolitics}

\input{chapters-en/01-topography}
\input{chapters-en/02-geometry}
\input{chapters-en/03-unified-equation}

%% ═══════════════════════════════════════════════════════════════════════════
%% PART II — The Unified DEQ² Apparatus
%% ═══════════════════════════════════════════════════════════════════════════
\part{The Unified \texorpdfstring{$\deq$}{DEQ²} Apparatus}

\input{chapters-en/04-systemic-form}
\input{chapters-en/05-four-variables}

%% ═══════════════════════════════════════════════════════════════════════════
%% PART III — The Four Actors
%% ═══════════════════════════════════════════════════════════════════════════
\part{The Four Actors}

\input{chapters-en/06-soliton-musk}
\input{chapters-en/07-decoherent-hegemon}
\input{chapters-en/08-rigid-synchronized}
\input{chapters-en/09-quantum-balance-brazil}

%% ═══════════════════════════════════════════════════════════════════════════
%% PART IV — The Phase Transition 2028--2031
%% ═══════════════════════════════════════════════════════════════════════════
\part[The Phase Transition 2028--2031]{The Phase Transition\\2028--2031}

\input{chapters-en/10-spc-dashboard}
\input{chapters-en/11-dated-predictions}

%% ═══════════════════════════════════════════════════════════════════════════
%% PART V — Antifragile Protocols
%% ═══════════════════════════════════════════════════════════════════════════
\part{Antifragile Protocols}

\input{chapters-en/12-antifragile-strategies}
\input{chapters-en/13-conclusions}

%% ═══════════════════════════════════════════════════════════════════════════
%% APPENDICES
%% ═══════════════════════════════════════════════════════════════════════════
\appendix

\input{chapters-en/AppA-formal-derivation}
\input{chapters-en/AppB-sigma-calibration}
\input{chapters-en/AppC-measurement-protocols}
\input{chapters-en/AppD-dialogue-traditions}

\end{document}
