\documentclass[12pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{geometry}
\geometry{a4paper, margin=2.5cm}

\address{Luis Fernando Marcelino\\
Independent Researcher\\
Milan, Italy\\
\texttt{lfrmrc@gmail.com}}

\date{\today}

\begin{document}

\begin{letter}{Editorial Office\\
\textit{Complexity}\\
Wiley / Hindawi}

\opening{Dear Editor,}

I am pleased to submit for your consideration the manuscript:

\medskip
\noindent\textbf{``From Discursive Topography to Hegemonic Phase Coherence: An Empirical Bridge via Kuramoto Synchronization''}
\medskip

This paper develops and empirically calibrates a formal multi-scale framework connecting communicative-microscopic dynamics to geopolitical-macroscopic stability. The key contributions are:

\begin{enumerate}
\item A proof of scale invariance in social-system payoff functionals: the cooperation threshold $\delta^* = 0.50$ derived independently from a communicative Folk Theorem and a hegemonic Folk Theorem are shown to be algebraically identical — the same mathematical structure at two scales.

\item A linear aggregation map $\Sigma^{\text{geo}} = J \Sigma^{\text{comm}} J^T$ that propagates micro-discursive covariances to macro-hegemonic covariances, calibrated on 52 nations (2014--2024) with bootstrap confidence intervals. The diagonal-scaling model (Strategy B) yields $R^2 = 0.84$ and is selected over the identity baseline ($p < 0.001$, LRT) and the full-matrix model (no improvement, $p = 0.53$).

\item A systemic stability functional $T_s^{\text{sys}}$ (Eq.~5) derived from Kuramoto's order parameter $\Phi$, with an asymmetric coupling between antifragility ($A$, multiplied by $\Phi$) and involutive dissipation ($\Omega$, divided by $\Phi$) — a mathematical analog of the second law applied to social systems.

\item Three publicly pre-registered falsifiable predictions for 2027--2031 (\texttt{github.com/bilanciere-quantistico/predizioni-deq2}).
\end{enumerate}

The manuscript is original and not under consideration elsewhere. An earlier preprint version appeared in Italian as part of the book \textit{Il Bilanciere Quantistico} (Marcelino, 2026, KDP); the present paper is a substantially expanded, reformulated, and independently verifiable scientific contribution.

The paper will be of interest to \textit{Complexity}'s readers working on social-system dynamics, multi-scale formalism, opinion dynamics, and geopolitical modeling. It applies Kuramoto synchronization --- a familiar tool in physical complex systems --- to a new domain (inter-hegemonic dynamics) with explicit empirical calibration, pre-registered predictions, and falsification conditions.

I have no conflicts of interest to declare. The data and code are publicly available at the pre-registration repository.

\closing{Sincerely,}

\end{letter}
\end{document}
